A cascaded supervised learning approach to inverse reinforcement learning

### 2013-03-26 piot2012cascading/ECML13_cascading.tex
Update piot2012cascading for ECML submission.

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without explicitely looking for the reward function. The reward function nevertheless is of interest in its own right. As mentionned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling between understanding the task and learning how to fulfill it that exists in AL. IRL allows the use of (A)DP or RL techniques to learn the task from the computed reward function. Although some approaches use it as a byproduct, most AL algorithm makes no use of the reward function, or have nothing to do with IRL altogether. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

Having in mind that there is a one to one relation between a reward function and its associated optimal action-value function (via the Bellman equation, see Eq.~\eqref{eq:bellman1}), it is worth thinking of a method able to output an action-value function for which the greedy policy is the demonstrated expert policy. Thus, the demonstrated expert policy will be optimal for the related reward function. We propose to use a score function-based multi-class classification step (see Sec.~\ref{sec:algo}) to infer a score function. Besides, in order to retrieve via the Bellman equation the reward associated to the score function computed by the classification step, we introduce a regression step (see Sec.~\ref{sec:algo}). That is why the method is called the Cascaded Supervised Inverse reinforcement learning (CSI). This method is analyzed in Sec.~\ref{sec:analysis}, where it is shown that the demonstrated expert policy is near-optimal for the reward the regression step outputs.

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-experts policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsec:instanciation}) is tested on an inverted pendulum simulator (Sec.~\ref{subsec:IP}) and a car driving simulator (Sec.~\ref{subsec:highway}) where we compare it with an AL algorithm \cite{abbeel2004apprenticeship}, with a pure classification algorithm \cite{taskar2005learning} and with a recent successfull IRL method \cite{klein2012scirl} as well as with a random baseline.

Differences and similarities with existing AL or IRL approaches are succinctly exposed in Sec.~\ref{sec:related}.
\section{Background and notations}
\label{sec:background}
First, we introduce some general notations.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
We note $\Delta_X$ the set of distributions over $X$.
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\geq\beta \Leftrightarrow \forall x\in X, \alpha(x) \geq \beta(x)$. We will often slightly abuse the notations and consider (where applicable) most objects as if they were matrices and vectors indexed by the set they operate upon.

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in ]0,1[$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. The stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies
\begin{equation}
  \rho_\pi^TP_\pi = \rho_\pi^T,
  \end{equation}
where $P_\pi$ is $(P(s'|s,\pi(s)))_{s,s' \in S}$. The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

@@

When the state space is too large to allow matricial representations or when the transition probabilites or even the reward function are unknown except through observations gained by interacting with the system, RL or ADP may be used to approximate the optimal control policy~\cite{sutton1998reinforcement}.

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although lots of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given label $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the label(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsec:instanciation}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation Eq.~\eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to solve the general, or approximate, IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitionned to. Action $a_j$ needs not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s=s_j,a=a_j)$. A regressor (a simple least-square approximator can do but other solutions such as Gaussian Processes \cite{rasmussen2006gaussian} could also be envisionned) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.
\begin{algorithm}%[H]
    %\small
  \caption{CSI algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi_E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{j},a_{j},s'_{j})\}_{1\leq j \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{((s_{j},a_{j}),\hat{r}_j)\}_{1\leq j \leq D'}$, $\forall (s_j,a_j,s'_j) \in D_R,\hat{r}_j=q(s_{j},a_{j})-\gamma q(s'_{j},\pi_C(s'_{j}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
\begin{equation}  
\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}];
\end{equation}
\item the regression error defined as 
  \begin{equation}
    \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E},
  \end{equation}
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are repectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$ ;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ could be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). We could upper bound $C_*$ by a more usual and general concentration coefficient but $C_*$ gives a tighter final result (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q$, defined by 
  \begin{eqnarray}
    \Delta q &= &\max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a)),
    \end{eqnarray} which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}

\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's introduce some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$. We recall the definition of the reward $R^C\in\mathbb{R}^{S\times A}$:
\begin{equation}
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s')),
\end{equation}
and the definition of $\hat{R}^C\in\mathbb{R}^{S\times A}$ which is the reward function output by the regression step.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data and a precise description of the way we instanciated the CSI algorithm are given on Sec.~\ref{subsec:instanciation}. 
We then use a classical RL benchmark, the Inverted Pendulum (Sec.~\ref{subsec:IP}), to provide a qualitative illustration of the proposed CSI approach.
We finally provide quantitative results and comparison with state-of-the-art approaches on a highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Instanciation}
\label{subsec:instanciation}
\subsubsection{Data scarcity}
\label{subsubsec:data}
In order to create a somewhat realistic environment, we decided that the CSI algorithm could neither use the transition probabilities nor a simulator. The only data it could access is sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

More precisely, the expert controls the system for $n$ runs of length $n$, with values of $n$ in $\{3,7,10,15,20\}$, giving $n^2$ samples $(s_k,a_k=\pi_E(a_k),s'_k)_k = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
\begin{equation}
  D_C = \{(s_i = s_k,a_i=a_k)_i\}.
\end{equation}
There is no particuliar problem for the classification step. It is however not reasonable to construct the dataset $D_R$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. The regression step also needs samples $((s_j,a_j),\hat r_j)_j$ drawn from non expert actions. We address this problem by using a heuristics.
\subsubsection{Heuristics}
\label{subsubsec:heuristics}
From the raw expert dataset $D_E$ we can construct samples $((s_j=s_k,a_j=a_k),\hat r_j)$ with $\hat r_j$ as specified in Eq.~\eqref{eq:rj}. To compensate for the lack of information about the dynamics of the whole MDP, we assert that disagreeing with the expert's choice is undesirable. Mathematically, we do that by artificially introducting samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
\begin{equation}
  r_{min} = \min_j\hat r_j - 1.
\end{equation}
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Subroutines choices}
\label{subsubsec:subroutines}
The cascading algorithm can be instanciated with some standard classification algorithms and any regression algorithm. The choice of such subroutines may be dictated by the kind and amount of available data, by ease of use or by computational complexity, for example.

>>>>

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without explicitely looking for the reward function. The reward function nevertheless is of interest in its own right. As mentionned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling that exists in AL between understanding the task and learning how to fulfill it. IRL allows the use of (A)DP or RL techniques to learn the task from the computed reward function. Although some approaches use it as a byproduct, most AL algorithm makes no use of the reward function, or have nothing to do with IRL altogether. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

Having in mind that there is a one to one relation between a reward function and its associated optimal action-value function (via the Bellman equation, see Eq.~\eqref{eq:bellman1}), it is worth thinking of a method able to output an action-value function for which the greedy policy is the demonstrated expert policy. Thus, the demonstrated expert policy will be optimal for the corresponding reward function. We propose to use a score function-based multi-class classification step (see Sec.~\ref{sec:algo}) to infer a score function. Besides, in order to retrieve via the Bellman equation the reward associated with the score function computed by the classification step, we introduce a regression step (see Sec.~\ref{sec:algo}). That is why the method is called the Cascaded Supervised Inverse reinforcement learning (CSI). This method is analyzed in Sec.~\ref{sec:analysis}, where it is shown that the demonstrated expert policy is near-optimal for the reward the regression step outputs.

This algorithm does not need to iteratively solve an MDP and needs only sampled transitions from expert and non-expert policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsec:instanciation}) is tested on an inverted pendulum simulator (Sec.~\ref{subsec:IP}) and a car driving simulator (Sec.~\ref{subsec:highway}) where we compare it with an AL algorithm \cite{abbeel2004apprenticeship}, with a pure classification algorithm \cite{taskar2005learning} and with a recent successfull IRL method \cite{klein2012scirl} as well as with a random baseline.

Differences and similarities with existing AL or IRL approaches are succinctly discussed in Sec.~\ref{sec:related}.
\section{Background and notations}
\label{sec:background}
First, we introduce some general notations.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
We note $\Delta_X$ the set of distributions over $X$.
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\geq\beta \Leftrightarrow \forall x\in X, \alpha(x) \geq \beta(x)$. We will often slightly abuse the notations and consider (where applicable) most objects as if they were matrices and vectors indexed by the set they operate upon.

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi$ is $(P(s'|s,\pi(s)))_{s,s' \in S}$, the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies
\begin{equation}
  \rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X.
  \end{equation}
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

@@

When the state space is too large to allow matrix representations or when the transition probabilites or even the reward function are unknown except through observations gained by interacting with the system, RL or ADP may be used to approximate the optimal control policy~\cite{sutton1998reinforcement}.

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although a lot of progress have been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given action\footnote{Here, actions play the role of what is known as {\it labels} or {\it categories} when talking about classifiers.} $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the action(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsec:instanciation}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation Eq.~\eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to solve the general, or approximate, IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s=s_j,a=a_j)$. A regressor (a simple least-square approximator can do but other solutions such as Gaussian Processes \cite{rasmussen2006gaussian} could also be envisionned) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

There is no particular constraint on $D_C$ and $D_R$. Evidently, there is a direct link between various qualities of those two sets (amount of data, statistical representativity, etc.) and the classification and regression error. The exact nature of the relationship between these quantities depends on which classifier and regressor are chosen. The theoretical analysis of Sec.\ref{sec:analysis} abstracts itself from the choice of a regressor and a classifier and from the composition of $D_C$ and $D_R$ by reasoning with the classification and regression error. In Sec.\ref{sec:experiments}, %FIXME subsec heuristic
the use of a single dataset to create both $D_C$ and $D_R$ is thoroughly explored.
\begin{algorithm}%[H]
    %\small
  \caption{CSI algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi_E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{j},a_{j},s'_{j})\}_{1\leq j \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{((s_{j},a_{j}),\hat{r}_j)\}_{1\leq j \leq D'}$, $\forall (s_j,a_j,s'_j) \in D_R,\hat{r}_j=q(s_{j},a_{j})-\gamma q(s'_{j},\pi_C(s'_{j}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
\begin{equation}  
\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}];
\end{equation}
\item the regression error defined as 
  \begin{equation}
    \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E},
  \end{equation}
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper bound by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions indeuced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q$, defined by 
  \begin{eqnarray}
    \Delta q &= &\max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a)),
    \end{eqnarray} which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
Readers familiar with the work presented in \cite{klein2012scirl} will see some similarities between the theoretical analyses of SCIRL and CSI. As both algorithms solve the same problem, both theorems naturally bound the same criterion. Another shared feature is the use of the score function $q$ of the classifier as a proxy for the action-value function of the expert. The same structure can therefore be found in both analyses as there is only so many ways to study error propagation in IRL algorithms.\\
The attentive reader, however, will percieve that similarities stop there. The error terms occurring in the two bounds are not related to one another. As CSI makes no use of the feature expectation of the expert, what is known as $\bar\epsilon_Q$ in \cite{klein2012scirl} does not appear in this analysis. Likewise, the regression error $\epsilon_R$ of this paper does not appear in the analysis of SCIRL, which does not use a regressor. Perhaps more subtly, the classification error and classification policy known in both papers as $\epsilon_C$ and $\pi_C$ are not the same. The classification policy of SCIRL is no tantamount to what is called $\pi_C$ here. For SCIRL, $\pi_C$ is the greedy policy for an approximation of the value function of the expert with respect to the reward output by the algorithm. For CSI, $\pi_C$ is the decision rule of the classifier, an object that is not aware of the structure of the MDP.
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's introduce some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$. We recall the definition of the reward $R^C\in\mathbb{R}^{S\times A}$:
\begin{equation}
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s')),
\end{equation}
and the definition of $\hat{R}^C\in\mathbb{R}^{S\times A}$ which is the reward function output by the regression step.

@@

%% Feel good comment
\section{Experiments}
\label{sec:experiments}
In this section, we put our approach under stress to empirically demonstrate its behavior. We begin by providing information pertaining to both benchmarks: an explanation about the amount and source of the available data, the rationale behing the heuristics we use to compensate for the dire data scarcity and a quick word about the contender CSI is compared to are given Sec.~\ref{subsec:generalities}. We supply quantitative results and comparison of CSI with state-of-the art approaches on first a classical RL benchmark (the mountain car) %FIXME: subsec moutnain car
and then on a more difficult highway driving simulator (Sec.~\ref{subsec:highway}).
\subsection{Generalities}
\label{subsec:generalities}
\subsubsection{Data scarcity}
\label{subsubsec:data}
The CSI algorithm was designed not to repeatedly solve the RL problem. This feature makes it particularly well-suited to environments where sampling the MDP is difficult or costly. In the experiments, CSIis fed only with data sampled according to the expert policy. This corresponds for example to a situation where a costly system can only be controlled by a trained operator as a bad control sequence could lead to a system breakdown.

More precisely, the expert controls the system for $M$ runs of lengths ${L_i}_{1\leq i \leq M}$, %with values of $n$ in $\{3,7,10,15,20\}$,
giving samples $(s_k,a_k=\pi_E(a_k),s'_k)_k = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
\begin{equation}
  D_C = \{(s_i = s_k,a_i=a_k)_i\}.
\end{equation}
\subsubsection{Heuristics}
\label{subsubsec:heuristics}

There is no particuliar problem for the classification step. It is however not reasonable to construct the dataset $D_R = \{((s_k,a_k=\pi_E(s_k)),\hat r_k)_k\}$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. 
This means that for a certain state $s_k$ we only know the corresponding expert action $a_k=\pi_E(s_k)$ and the following state $s'_k$ sampled according to the MDP dynamics : $s'_k \sim P(\cdot|s_k,a_k)$. For the regression to be meaningful, we need samples associating the same state $s_k$ and a different action $a\neq a_k$ with a datapoint $\hat r \neq \hat r_k$.\\
Recall that $\{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j$ (Eq.~\ref{eq:rj}), without knowing $s'_k \sim P(\cdot|s_k,a\neq a_k)$, we can not provide the regressor with a datapoint to associate with $(s_k,a\neq a_k)$.
We will artificially augment the dataset $D_R$ with samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
\begin{equation}
  r_{min} = \min_j\hat r_j - 1.
\end{equation}
This heuristics intructs the regressor to associate a state-action tuple disagreeing with the expert ({\it i.e. }$(s_k,a\neq a_k)$) with a reward strictly inferior to any of those associated with expert state action tuples ({\it i.e. }$(s_k,a_k=\pi_E(s_k))$). Semantically, we are asserting that disagreeing with the expert in states the expert visited is a bad idea, thus the shape of the reward function should reflect it. This heuristics says nothing about states absent from the expert dataset. For such states the generalization capabilities of the regressor and, later on, the exploration of the MDP by an agent optimizing the reward will solve the problem.
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Comparison with state of the art approaches}
The similar looking yet fundamentally different algorithm SCIRL is an obvious choice as a contender to CSI as it advertises the same ability to work with very little data, without repeatedly solving the RL problem. In both experiments we give the exact same data to CSI and SCIRL.
